Applying the metallicity-dependent binary fraction to double white dwarf formation: \\ Implications for LISA

### 2021-11-26 src/ms.tex
Update ms.tex

When viewed strictly as a source of noise, the unresolved Galactic DWD foreground is the dominant noise source for LISA in the sub-mHz part of LISA's frequency range. This extra noise above the detector noise floor affects the detection of all other LISA sources including extreme mass ratio inspirals \citep[e.g.,][]{Berti2006, Barack2007, Babak2017, Moore2017}, merging black holes with masses between $10^4$--$10^7\,M_{\odot}$ \citep[e.g.,][]{Klein2016, Bellovary2019}, and cosmological GW backgrounds \citep[e.g.,][]{Bartolo2016, Caprini2016, Caldwell2019}. For sources which have signals buried by the Galactic DWD population, the foreground must be carefully analyzed and subtracted \citep[][]{Adams2014,Cornish2020,Littenberg2020,Boileau2021}. The number of resolved DWDs and the height of the unresolved DWD foreground are a direct consequence of the number of DWD progenitors which form and evolve over the Milky Way's history.

@@

The masses and orbital periods at the formation of the second-formed WD span a wide range depending on the WD binary component types, thus we consider four DWD combinations: two helium WDs (He + He), a carbon-oxygen WD orbiting a helium WDs (CO + He), two carbon-oxygen WDs (CO + CO), and an oxygen-neon orbiting helium, carbon-oxygen, or oxygen-neon WD (ONe + X). For each DWD type we simulate a grid of $15$ metallicities spaced uniformly in $\log_{10}(Z)$ between $Z=10^{-4}$ to $0.03$, to account for the limits of the \citet{Hurley2000} stellar evolution tracks employed in \cosmic. This results in a total of $60$ populations across all DWD types and metallicities. The output of \cosmic\ contains information limited to intrinsic binary properties like mass and orbital period. External parameters like Galactic position and orientation are assigned in a post-processing scheme which uses metallicity-dependent positions and ages from the Ananke framework of galaxy \textbf{m12i} from the Latte Suite of the FIRE simulations (see Section~\ref{sec:FIRE} for details).

@@

\subsection{A metallicity-dependent SFH: Convolving with the FIRE models}
\label{sec:FIRE}
To create Milky Way-like galaxies which integrate the metallicity-dependent binary fraction, we use the metallicity-dependent ages and positions of galaxy \textbf{m12i} from the ``Latte" suite of the FIRE simulations \citep{Hopkins2015, Wetzel2016, Hopkins2018} to create synthetic, Milky Way-like DWD populations.

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{figures/sfh_vs_fb.pdf}
    \caption{The metallicity-dependent binary fraction for close solar-type binaries with P $< 10^4$ days (black) plotted against the logarithm of Z/Z$_\odot$, with scatter points denoting the location of the metallicity grid used in our \cosmic\ simulations. The secondary axis shows the amount of mass in stars formed within each metallicity bin from galaxy \textbf{m12i} of the ``Latte" suite in FIRE-2 simulations as a red histogram. The amount of stellar mass formed with super-solar metallicities dominates the distribution. Note that the primary axis shows $f_{\rm{b}}(Z)$ in linear scale, and the secondary axis shows amount of mass formed in log-scale. The opposing trends of these two distributions compete throughout this study.}
    \label{fig:sfh_vs_fb}
\end{figure}

@@

\begin{equation}
    a_{\text{RLO,}\ell} = R_{\ell} \frac{0.6 q_{\ell}^{2/3} + \ln{(1+q_{\ell}^{1/3})}}{0.49 q_{\ell}^{2/3}}
\end{equation}
where $R_{\ell}$ is the radius of the lower mass WD and $q_{\ell} = M_{\ell}/M_{h}$ is the ratio of the lower- to higher-mass WD components. \citep{Eggleton1983}. 
We define the radius of a WD following \citet{Tout1997, Hurley2000} as
\begin{equation}
    R_{\rm{WD}} = \max\Bigg(R_\text{NS}, 0.0115\sqrt{\left(\frac{M_\text{Ch}}{M}\right)^{2/3}-\left(\frac{M}{M_\text{Ch}}\right)^{2/3}} \Bigg) 
\end{equation}
\noindent where $R_\text{NS} = 1.4\cdot 10^{-5}$ R$_\odot$ is the radius of a neutron star, $M_{\text{Ch}}=1.44$ M$_\odot$ is the Chandrasekhar limit for the mass of a stable WD, and $M$ is the mass of the DWD in solar masses.

@@

For stationary sources, the SNR is modified due to the lack of observable orbital evolution as
\begin{equation}
    \rho_{\text{circ,stat}}^2 = \frac{h_{2}^2T_{\text{obs}}}{4 S_n(f_2)}
\end{equation}
with the observation time $T_{\text{obs}}$. Here, $h_2$ is strain amplitude of the source for the second orbital frequency harmonic
\begin{equation}
   h_2^2 = \frac{2^{22/3}}{5}\frac{(G\mathcal{M}_c)^{10/3}}{c^8D_L^2}(\pi f_\text{orb})^{4/3}.
\end{equation}
and is connected to the characteristic strain as
\begin{equation}
    h_2^2 = \frac{\dot{f}_{2}}{f_{\rm{orb}}^2} h_c^2.
\end{equation}

@@

If a lower-metallicity system forms a CO + He DWD and a higher-metallicity system does not, this is because the initially more massive binary component initiates a CE while on the giant branch (GB) instead of the AGB. This instead leaves behind a He WD with a stellar companion, which is the dominating scenario that restricts formation efficiency. There are a few edge cases where either a He + He DWD is formed instead, or when a CO WD and a stellar companion is formed and there has been very nearly but not quite enough time for a CO + He DWD to form.

>>>>

When viewed strictly as a source of noise, the unresolved Galactic DWD foreground is the dominant noise source for LISA in the sub-mHz part of LISA's frequency range. This extra noise above the detector noise floor affects the detection of all other LISA sources including extreme mass ratio inspirals \citep[e.g.,][]{Berti2006, Barack2007, Babak2017, Moore2017}, merging black holes with masses between $10^4$--$10^7$ M$_{\odot}$ \citep[e.g.,][]{Klein2016, Bellovary2019}, and cosmological GW backgrounds \citep[e.g.,][]{Bartolo2016, Caprini2016, Caldwell2019}. For sources which have signals buried by the Galactic DWD population, the foreground must be carefully analyzed and subtracted \citep[][]{Adams2014,Cornish2020,Littenberg2020,Boileau2021}. The number of resolved DWDs and the height of the unresolved DWD foreground are a direct consequence of the number of DWD progenitors which form and evolve over the Milky Way's history.

@@

The masses and orbital periods at the formation of the second-formed WD span a wide range depending on the WD binary component types, thus we consider four DWD combinations: two helium WDs (He + He), a carbon-oxygen WD orbiting a helium WD (CO + He), two carbon-oxygen WDs (CO + CO), and an oxygen-neon orbiting a helium, carbon-oxygen, or oxygen-neon WD (ONe + X). For each DWD type we simulate a grid of $15$ metallicities spaced uniformly in $\log_{10}(Z)$ between $Z=10^{-4}$ to $0.03$, to account for the limits of the \citet{Hurley2000} stellar evolution tracks employed in \cosmic. This results in a total of $60$ populations across all DWD types and metallicities. The output of \cosmic\ contains information limited to intrinsic binary properties like mass and orbital period. External parameters like Galactic position and orientation are assigned in a post-processing scheme which uses metallicity-dependent positions and ages from the Ananke framework of galaxy \textbf{m12i} from the Latte Suite of the FIRE-2 simulations (see Section~\ref{sec:FIRE} for details).

@@

\subsection{A metallicity-dependent SFH: Convolving with the FIRE-2 models}
\label{sec:FIRE}
To create Milky Way-like galaxies which integrate the metallicity-dependent binary fraction, we use the metallicity-dependent ages and positions of galaxy \textbf{m12i} from the ``Latte" suite of the FIRE-2 simulations \citep{Hopkins2015, Wetzel2016, Hopkins2018} to create synthetic, Milky Way-like DWD populations.

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{figures/sfh_vs_fb.pdf}
    \caption{The metallicity-dependent binary fraction for close solar-type binaries with P $< 10^4$ days (black) plotted against the logarithm of Z/Z$_\odot$, with scatter points denoting the location of the metallicity grid used in our \cosmic\ simulations. The secondary axis shows the amount of mass in stars formed within each metallicity bin from galaxy \textbf{m12i} of the ``Latte" suite in the FIRE-2 simulations as a red histogram. The amount of stellar mass formed with super-solar metallicities dominates the distribution. Note that the primary axis shows $f_{\rm{b}}(Z)$ in linear scale, and the secondary axis shows amount of mass formed in log-scale. The opposing trends of these two distributions compete throughout this study.}
    \label{fig:sfh_vs_fb}
\end{figure}

@@

\begin{equation}
    a_{\text{RLO,}\ell} = R_{\ell} \frac{0.6 q_{\ell}^{2/3} + \ln{(1+q_{\ell}^{1/3})}}{0.49 q_{\ell}^{2/3}}
\end{equation}
where $R_{\ell}$ is the radius of the lower mass WD and $q_{\ell} = M_{\ell}/M_{h}$ is the ratio of the lower- to higher-mass WD components. \citep{Eggleton1983}. 
We define the radius of a WD following \citet{Tout1997, Hurley2000} as
\begin{equation}
    R_{\rm{WD}} = \max\Bigg(R_\text{NS}, 0.0115\sqrt{\left(\frac{M_\text{Ch}}{M}\right)^{2/3}-\left(\frac{M}{M_\text{Ch}}\right)^{2/3}} \Bigg) 
\end{equation}
\noindent where $R_\text{NS} = 1.4\cdot 10^{-5}$ R$_\odot$ is the radius of a neutron star, $M_{\text{Ch}}=1.44$ M$_\odot$ is the Chandrasekhar limit for the mass of a stable WD, and $M$ is the mass of the WD in solar masses.

@@

For stationary sources, the SNR is modified due to the lack of observable orbital evolution as
\begin{equation}
    \rho_{\text{circ,stat}}^2 = \frac{h_{2}^2T_{\text{obs}}}{4 S_n(f_2)}
\end{equation}
with the observation time $T_{\text{obs}}$. Here, $h_2$ is strain amplitude of the source for the second orbital frequency harmonic,
\begin{equation}
   h_2^2 = \frac{2^{22/3}}{5}\frac{(G\mathcal{M}_c)^{10/3}}{c^8D_L^2}(\pi f_\text{orb})^{4/3}
\end{equation}
and is connected to the characteristic strain as
\begin{equation}
    h_2^2 = \frac{\dot{f}_{2}}{f_{\rm{orb}}^2} h_c^2
\end{equation}

@@

If a lower-metallicity system forms a CO + He DWD and a higher-metallicity system does not, this is because the initially more massive binary component initiates a CE while on the giant branch (GB) instead of the asymptotic giant branch (AGB). This instead leaves behind a He WD with a stellar companion, which is the dominating scenario that restricts formation efficiency. There are a few edge cases where either a He + He DWD is formed instead, or when a CO WD and a stellar companion is formed and there has been very nearly but not quite enough time for a CO + He DWD to form.
